Erosion of somatic tissue identity with loss of the X-linked intellectual disability factor KDM5C

### 2024-08-30 submission/manuscript.Rmd
add log2fc to 2-cell genes

zscan4d <- epi_restab[epi_restab$ENSEMBL == 'ENSMUSG00000090714',]['padj'][,1]
duxf3 <- epi_restab[epi_restab$ENSEMBL == 'ENSMUSG00000075046',]['padj'][,1]

@@

Notably, DEGs unique to _Kdm5c_-KO EpiLCs included key drivers of germline identity, such as _Stimulated by retinoic acid 8_ (_Stra8_: log2 fold change = `r signif(getresval(epigerm, 'log2FoldChange', 'Stra8'), 3)`, q = `r signif(getresval(epigerm, 'padj', 'Stra8'), 3)`) and _Deleted in azoospermia like_ (_Dazl_: log2 fold change = `r signif(getresval(epigerm, 'log2FoldChange', 'Dazl'), 3)`, q = `r signif(getresval(epigerm, 'padj', 'Dazl'), 3)`) (Figure 3H). These genes are typically expressed when primordial germ cells (PGCs) are committed to the germline fate, but are also expressed later in life to trigger meiotic gene expression programs[@koubovaRetinoicAcidRegulates2006; @linGermCellIntrinsicExtrinsic2008; @endoRetinoicAcidGerm2019]. Of note, some germline genes, including _Dazl_, are also expressed in the two-cell embryo[@macfarlanEmbryonicStemCell2012; @guptaGenomewideScreenReveals2023]. However, we did not see derepression of two-cell stage-specific genes, like _Duxf3 (Dux)_ (q = `r signif(duxf3, 3)`) and _Zscan4d_ (q = `r signif(zscan4d, 3)`) (Figure 3H), indicating _Kdm5c_-KO EpiLCs do not revert back to a 2-cell-like state. Altogether, _Kdm5c_-KO EpiLCs expressing key drivers of germline identity and meiosis while the brain primarily expresses spermiogenesis genes indicate germline gene misexpression parallels germline development during proper _Kdm5c_-KO differentiation.

>>>>

zscan4d_padj <- epi_restab[epi_restab$ENSEMBL == 'ENSMUSG00000090714',]['padj'][,1]
duxf3_padj <- epi_restab[epi_restab$ENSEMBL == 'ENSMUSG00000075046',]['padj'][,1]

zscan4d_l2fc <- epi_restab[epi_restab$ENSEMBL == 'ENSMUSG00000090714',]['log2FoldChange'][,1]
duxf3_l2fc <- epi_restab[epi_restab$ENSEMBL == 'ENSMUSG00000075046',]['log2FoldChange'][,1]

@@

Notably, DEGs unique to _Kdm5c_-KO EpiLCs included key drivers of germline identity, such as _Stimulated by retinoic acid 8_ (_Stra8_: log2 fold change = `r signif(getresval(epigerm, 'log2FoldChange', 'Stra8'), 3)`, q = `r signif(getresval(epigerm, 'padj', 'Stra8'), 3)`) and _Deleted in azoospermia like_ (_Dazl_: log2 fold change = `r signif(getresval(epigerm, 'log2FoldChange', 'Dazl'), 3)`, q = `r signif(getresval(epigerm, 'padj', 'Dazl'), 3)`) (Figure 3H). These genes are typically expressed when primordial germ cells (PGCs) are committed to the germline fate, but are also expressed later in life to trigger meiotic gene expression programs[@koubovaRetinoicAcidRegulates2006; @linGermCellIntrinsicExtrinsic2008; @endoRetinoicAcidGerm2019]. Of note, some germline genes, including _Dazl_, are also expressed in the two-cell embryo[@macfarlanEmbryonicStemCell2012; @guptaGenomewideScreenReveals2023]. However, we did not see derepression of two-cell stage-specific genes, like _Duxf3 (Dux)_ (log2 fold change = `r signif(duxf3_l2fc, 3)`, q = `r signif(duxf3_padj, 3)`) and _Zscan4d_ (log2 fold change =  `r signif(zscan4d_l2fc, 3)`, q = `r signif(zscan4d_padj, 3)`) (Figure 3H), indicating _Kdm5c_-KO EpiLCs do not revert back to a 2-cell-like state. Altogether, _Kdm5c_-KO EpiLCs expressing key drivers of germline identity and meiosis while the brain primarily expresses spermiogenesis genes indicate germline gene misexpression parallels germline development during proper _Kdm5c_-KO differentiation.
